---
title: "Figure 4A: Survival Progeny experiment"
date: "`r Sys.Date()`"
output: html_document
---
# Experimental details
 Virus:     Progeny DCV from flies with(W+) or without wolbachia (W-)
 Titre:     50.6 nL of 10E12 total virus particles/mL 
 Fly:       w1118 wMelPopTet
 Figure:    4B
 
# Rationale 
To determine if Wolbachia has an impact on the ability for progeny virus(W- or W+) to cause disease in Drosophila melanogaster host. 

# Approach 
Use the coxme model. 
 
# Preparation 
```{r Load librarys}
# Load library 
library(coxme)
library(survminer)
library(tidyverse)
```

# Importing data
Day         = Day at which mortality was seen else the last day of the experiment
Suvival     = Whether the fly was alive (0) or dead (1) on the day 
Virus       = Presence of virus (0 = absent, 1 = present)
WolVirus    = Type of progeny virus used (0 = W-, 1 = W+)
TechRep     = Individual vial a fly was part of 
BiolRep     = Virus prep used to infect the fly 

```{r Importing data}
# Importing data
data_DCV_progeny <- read.csv("Survival_Progeny_DCV_Figure_4_B_data.csv")

# Formatting data_DCV_progeny
data_DCV_progeny$WolVirus <- factor(data_DCV_progeny$WolVirus) # Wolvirus factor 
data_DCV_progeny$Virus <- factor(data_DCV_progeny$Virus) # Virus as factor 
data_DCV_progeny_noPBS <- dplyr::filter(data_DCV_progeny, Virus == 1) # No PBS
data_DCV_progeny_PBS <- dplyr::filter(data_DCV_progeny, Virus == 0) # No virus 

# Checking data_DCV_progeny 
head(data_DCV_progeny)
head(data_DCV_progeny_noPBS)
head(data_DCV_progeny_PBS)

# Checking factors coded correctly 
levels(data_DCV_progeny$WolVirus)
levels(data_DCV_progeny$Virus)
```

# Coxme Analysis 
Applying Cox-me model to determine if there is a difference between the surivival of flies infected with  W- virus when compared with those infected with W+ DCV
Model is fit with Virus type (WolVirus) as the predictor and biol and tech replication as random effects.

```{r Coxme fitting}
# Fitting the model 
fit1 <- coxme(Surv(Days, Survival) ~ WolVirus + (1 | BiolRep/TechRep), data = data_DCV_progeny_noPBS)
```

# Assumptions 
Assumptions of coxme: 
1. Proportional hazards 

Results:
1. Passed: p-value for global and wolvirus is not significant

Conclusions: 
As assumptions are meet conduct Cox-me and report analysis


```{r Coxme Assumptions}
# 1. Proportional hazards: Formally test
ftest <- cox.zph(fit1)
print(ftest) # prints test results
ggcoxzph(ftest) # plots test results
```

# Conclusions 
From the analysis and the summary below
 Wolbachia does NOT impact the pathogenicity of progeny virus
* The HR of flies infected with W+ virus is 0.8968857 (coxme) and as the CI crosses 1 there is no difference 
* For each time point there is no difference in the probability of death between flies infected with W- and W+ DCV virus 

# Summary of analysis 
```{r Summary Coxme output}
# Extract the HR estimates and confidence intervals
coef_cox<- as.data.frame(summary(fit1)$coefficients)
HR_cox <- as.data.frame(exp(coef_cox)) # Calculates hazard ratio (HR)
CI_cox <- as.data.frame(exp(confint(fit1))) # Calculates confidence interval

# Extracts p-value 
extract_coxme_table <- function (mod){
  beta <- mod$coefficients 
  nvar <- length(beta)
  nfrail <- nrow(mod$var) - nvar
  se <- sqrt(diag(mod$var)[nfrail + 1:nvar])
  z<- round(beta/se, 2)
  p<- signif(1 - pchisq((beta/se)^2, 1), 2)
  table=data.frame(cbind(beta,se,z,p))
  return(table)
}
extract_coxme_table(fit1)
p_cox <- as.data.frame(extract_coxme_table(fit1)$p)

# Creates a data frame with the cofidence intervals etc 
df_cox <- bind_cols(coef_cox, HR_cox, CI_cox, p_cox) %>% 
  rename(coef = `summary(fit1)$coefficients...1`, 
         HR = `summary(fit1)$coefficients...2`, 
         "LB_2.5%" = "2.5 %",
         "UB_97.5%" = "97.5 %", 
         pval = "extract_coxme_table(fit1)$p") %>% 
  mutate(test = "coxme") %>% 
  tibble::rownames_to_column("factor")

# confirms extraction 
print(fit1)
View(df_cox)
```

```{r Visualising Coxme output}
# Graph of HR and CI 
plot_coxme <- ggplot(df_cox, aes(x = factor, y = HR, colour = test, group_by(test))) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = `LB_2.5%`, ymax = `UB_97.5%`, color = test), width = 0.1) +
  coord_flip() +
  xlab("") +
  ylab("Hazard Ratio") +
  scale_color_manual(values = c("blue")) +
  theme_minimal()
```

# Graphing Kaplan-Meier survival curves 
The curves generated are Kaplan-Meier but the analysis that has been conducted is coxme

```{r Creating Kaplan-Meier curves}
# Creating the survival curve 
curves <- survfit(Surv(Days, Survival) ~ WolVirus, data=data_DCV_progeny_noPBS)
```

```{r Plotting Kaplan-Meier curves}
graph_DCV_progeny <- ggsurvplot(curves,
           data = data_DCV_progeny_noPBS,               #change to the virus under analy
           axes.offset = FALSE,            #changes axes to start at zero
           line = c(1,1,1,2,2,2),
           xlim = c(0,10),
           ylim = c(0,1.07),
           xlab = "Days Post Infection",
           ylab = "Proportional survival",
           legend.title=element_blank(), 
           break.time.by = 1, 
           break.y.by = 0.2,
           conf.int = TRUE,
           ggtheme = theme_classic2(),
           palette = c("black", "red"),
           legend = "none"
) |> ggpar(
  font.x        = c(30, "bold", "black"),          
  font.y        = c(30, "bold", "black"),      
  font.xtickslab = c(30, "plain", "black"),
  font.ytickslab = c(30, "plain", "black")
) 

```

# Exporting outputs 
```{r Exporting outputs}
# Saving coxme summary
write.csv(df_cox, file = "Survival_Progeny_DCV_Figure_4_B_output_coxme_HR.csv", row.names = FALSE)

# Saving coxme fit output 
sink("Survival_Progeny_DCV_Figure_4_B_output_coxme_summary.txt") # change file name
summary(fit1) # prints coxme output to txt file UP TO HERE
sink() # closes the connection **CRITICAL**


# Saving coxme plot 
ggsave("Survival_Progeny_DCV_Figure_4_B_output_coxme_plot.pdf", 
       plot = plot_coxme, 
       height = 7.5, 
       width = 15) 

# Saving Kaplan-Meier survival curves 
ggsave("Survival_Progeny_DCV_Figure_4_B_output_curves_plot.pdf", 
       plot = graph_DCV_progeny$plot, 
       height = 7.5, 
       width = 15)
```
