---
title: "Figure 1A: Survival DCV Original "
date: "`r Sys.Date()`"
output: html_document
---
# Experimental details
 Virus: DCV
 Titre: 10E8 IU/mL 
 Fly: w1118 wMelPop and w1118 wMelPopTet
 Figure: 1A
 
# Rationale 
To determine the impact Wolbachia has on the survival of flies infected with DCV. 

# Approach 
Use the coxme model
If doesn't mean proportional hazards 
* Non-parametric alternatives(Kutzner et al. 2023, Kutzner et al. 2019)

# Preparation 
```{r Load librarys}
# Load library 
library(coxme)
library(survminer)
library(SurvRegCensCov)
library(tidyverse)
```

# Importing data
Day         = Day at which mortality was seen else the last day of the experiment
Suvival     = Whether the fly was alive (0) or dead (1) on the day 
Wol         = Presence or absences of Wolbachia 0 = Absent w1118-wMelpopTet, 1 = present w1118-wMelPop)
Virus       = Presence of virus (0 = absent, 1 = present)
TechRep     = Individual vial a fly was part of 
BiolRep     = cohort that an individaul fly was a part of 


```{r Importing data}
data_DCV <- read.csv("Survival_DCV_Figure_1_A_data.csv")
```

```{r Formating data}
# Formatting data
data_DCV$Wol <- factor(data_DCV$Wol) # Wolbachia as factor
data_DCV$Virus <- factor(data_DCV$Virus) # Virus as factor
data_DCV_noPBS <- dplyr::filter(data_DCV, Virus == 1) # filters out PBS
data_DCV_PBS <- dplyr::filter(data_DCV, Virus == 0) # Filters out virus 

```

# Coxme Analysis 
Applying cox-me model to determine the impact of Wolbachia on survival when flies are infected with DCV.

* Wolbachia = predictor
* biol and tech replicates = random effects

```{r Coxme fitting}
# Fitting the model 
fit1 <- coxme(Surv(Days, Survival) ~ Wol + (1 | BiolRep/TechRep), data = data_DCV_noPBS)
```

# Assumptions 
Testing the assumptions of coxme
1. Independent of treatments, virus impacts survival 
2. Independent of viral infection, Wolbachia doesn't impact survival (PBS)
3. Proportional hazards 

Results:
1. Passed: p-value for virus coef is significant 
2. Passed: p-value for wol coef is not significant
3. Fail: Global and wol coef fails a p-value significant

Conclusions: 
Going forward proceed with the fit1 coxme and test using the non-parameteric analysis to confirm the results are qualitatively similar. 

```{r Coxme Assumptions}
# 1. Independent of treatments virus impacts survival 
fitSus <- coxme(Surv(Days, Survival) ~ Virus + (1 | BiolRep/TechRep), data= data_DCV) 
fitSus 

# 2. Independent of viral infection Wolbachia doesn't impact survival (PBS)
fitPBS <- coxme(Surv(Days, Survival) ~ Wol + (1 | BiolRep/TechRep), data = data_DCV_PBS) 
fitPBS

# 3. Proportional hazard: Visualise
# Plot
LMLplot <- plot(survfit(Surv(Days, Survival) ~ Wol, data = data_DCV_noPBS), fun = "cloglog")

# 3. Proportional hazards: Formally test
ftest <- cox.zph(fit1)
print(ftest) # prints test results
ggcoxzph(ftest) # plots test results
```

# Comparing parametric with non-parametric coxme 
The assumption (3) of proportional hazards was not met. 
Coxme may not be appropriate.
A parametric model with no assumptions was fit to compare. 

Approach: 
Determining if the two tests (coxme and a parametric test) output is comparable. 
If they are then the coxme results can be reported. 

Results: 
Weibull distribution is the best fit model 
1. CI of overlap

Conclusions: 
Both models show that wol has a significant impact on survival. 
Report the coxme model results 


```{r Parametric test}
# Coxme
summary(fit1)

# Parametric analysis (fitting different parametric models)
fit_wei <-
  survreg(Surv(Days, Survival) ~ Wol , dist = "weibull", data = data_DCV_noPBS)
fit_expo <-
  survreg(Surv(Days, Survival) ~ Wol , dist = "exponential", data = data_DCV_noPBS)
fit_gauss <-
  survreg(Surv(Days, Survival) ~ Wol , dist = "gaussian", data = data_DCV_noPBS)
fit_extreme <-
  survreg(Surv(Days, Survival) ~ Wol , dist = "extreme", data = data_DCV_noPBS)
fit_logit <-
  survreg(Surv(Days, Survival) ~ Wol , dist = "logistic", data = data_DCV_noPBS)

# Testing which model is the best
AIC(fit_wei, fit_expo, fit_gauss, fit_extreme, fit_logit) # Weibull was the best fit

```

```{r Comparing coxme with a parametric model}
# Summarising the best model
summary(fit_wei)
ConvertWeibull(fit_wei)
WeibullDiag(Surv(Days, Survival)~ Wol, data = data_DCV_noPBS)

# Extract coxme HR and CI 
coef_cox<- as.data.frame(summary(fit1)$coefficients)
HR_cox <- as.data.frame(exp(coef_cox))
CI_cox <- as.data.frame(exp(confint(fit1))) # coefficient CIs for HR 

# Extract weibull HR and CI 
coef_wei<- as.data.frame(ConvertWeibull(fit_wei)$vars[3,1])
HR_wei <- as.data.frame(ConvertWeibull(fit_wei)$HR) # Include upper and lower bounds

# Extracts p-value 
extract_coxme_table <- function (mod){
  beta <- mod$coefficients #$fixed is not needed
  nvar <- length(beta)
  nfrail <- nrow(mod$var) - nvar
  se <- sqrt(diag(mod$var)[nfrail + 1:nvar])
  z<- round(beta/se, 2)
  p<- signif(1 - pchisq((beta/se)^2, 1), 2)
  table=data.frame(cbind(beta,se,z,p))
  return(table)
}
extract_coxme_table(fit1)

# Storing p-values 
p_cox <- as.data.frame(extract_coxme_table(fit1)$p)
p_wei <- as.data.frame(extract_coxme_table(fit_wei)[2,]$p)

# Create data_DCVframes
df_cox <- bind_cols(coef_cox, HR_cox, CI_cox, p_cox) %>% 
  rename(coef = `summary(fit1)$coefficients...1`, 
         HR = `summary(fit1)$coefficients...2`, 
         "LB_2.5%" = "2.5 %",
         "UB_97.5%" = "97.5 %", 
         pval = "extract_coxme_table(fit1)$p") %>% 
  mutate(test = "coxme") %>% 
  tibble::rownames_to_column("factor")

df_wei <- bind_cols(coef_wei, HR_wei, p_wei)%>% 
  rename(coef = `ConvertWeibull(fit_wei)$vars[3, 1]`,
         "LB_2.5%" = "LB", 
         "UB_97.5%" = "UB", 
         pval = "extract_coxme_table(fit_wei)[2, ]$p") %>% 
  mutate(test = "weibull") %>% 
  tibble::rownames_to_column("factor")

# Summary df of all the information
df_join <- full_join(df_cox, df_wei, by = join_by(factor, coef, HR, `LB_2.5%`, `UB_97.5%`,pval, test))

# confirms extraction 
print(fit1)
print(fit_wei)
```

```{r Visualising coxme and Weibull output}
# Graph of HR and CI 
plot_all <- ggplot(df_join, aes(x = factor, y = HR, colour = test, group_by(test))) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = `LB_2.5%`, ymax = `UB_97.5%`, color = test), width = 0.1) +
  coord_flip() +
  xlab("") +
  ylab("Hazard Ratio") +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal()
```


# Conclusions 
From the analysis and the summary
HR of both the parametric and non-parametric analysis is close and the CI over lap.Therefore only reported the results from the coxme. 

Wolbachia impacts survival (coxme)
* The HR of flies with wol is 0.08902321
* For every fly without wolbachia 0.08902321 flies with Wolbachia die. 
* For each time point flies with wol are ~92% (1-0.08902321 = 0.92) less likely to die from virus than flies without 

# Graphing Kaplan-Meier survival curves 
The curves generated for visualisation purposes are Kaplan-Meier but the analysis that has been conducted is coxme

```{r Creating Kaplan-Meier curves}
curves <- survfit(Surv(Days, Survival) ~ Wol, data=data_DCV_noPBS)
```

```{r Plotting Kaplan-Meier curves}

graph_DCV <- ggsurvplot(curves,
           data = data_DCV_noPBS,         
           axes.offset = FALSE,            #changes axes to start at zero
           line = c(1,1,1,2,2,2),
           xlim = c(0,25),
           ylim = c(0,1.01),
           xlab = "Days Post Infection",
           ylab = "Proportional survival",
           legend.title=element_blank(), 
           break.time.by = 2, 
           break.y.by = 0.2,
           conf.int = TRUE,
           ggtheme = theme_classic2(),
           palette = c("black", "red"),
           legend = "none"
) |> ggpar(
  font.x        = c(30, "bold", "black"),          
  font.y        = c(30, "bold", "black"),      
  font.xtickslab = c(30, "plain", "black"),
  font.ytickslab = c(30, "plain", "black")
)
```


# Exporting outputs
```{r Exporting outputs - Run outside RDM}

# Saving coxme and weibull summary 
write.csv(df_join, file = "Survival_DCV_Figure_1_A_output_coxme_HR.csv", row.names = FALSE)

# Exporting stats output
sink("Survival_DCV_Figure_1_A_output_coxme_summary.txt") # change file name
summary(fit1) # prints coxme output to txt file UP TO HERE
sink() # closes the connection **CRITICAL**

# Saving coxme and weibull plot 
ggsave("Survival_DCV_Figure_1_A_output_plot_coxme.pdf", 
       plot = plot_all, 
       height = 7.5, 
       width = 15) 

# Saving Kaplan-Meier survival curves
ggsave("Survival_DCV_Figure_1_A_output_plot_survival_curves.pdf", 
       plot = graph_DCV$plot, 
       height = 7.5, 
       width = 15)
```