---
title: "Figure 2C: Survival FHV Original"
date: "`r Sys.Date()`"
output: html_document
---
# Experimental details
 Virus: FHV
 Titre: 10E9 IU/mL 
 Fly: w1118 wMelPop and w1118 wMelPopTet
 Figure: 2C
 
# Rationale 
To determine the impact of Wolbachia on the survival of flies infected with FHV 

# Approach 
Use the coxme model

# Resources 
Non-parametric alternatives 
 (Kutzner et al. 2023, Kutzner et al. 2019)

# Preparation 
```{r Load librarys}
# Load library 
library(coxme)
library(survminer)
library(SurvRegCensCov)
library(tidyverse)

```

# Importing data
Day         = Day at which mortality was seen else the last day of the experiment
Suvival     = Whether the fly was alive (0) or dead (1) on the day 
Wol         = Presence or absences of Wolbachia 0 = Absent w1118-wMelpopTet, 1 = present w1118-wMelPop)
Virus       = Presence of virus (0 = absent, 1 = present)
TechRep     = Individual vial a fly was part of 
BiolRep     = cohort that an individaul fly was a part of 

```{r Importing Data}
data_FHV <- read.csv("Survival_FHV_Figure_2_C_data.csv")
```

```{r Formating Data}
# Formatting data
data_FHV$Wol <- factor(data_FHV$Wol) # Wolbachia as factor
data_FHV$Virus <- factor(data_FHV$Virus) # Virus as factor
data_FHV_noPBS <- dplyr::filter(data_FHV, Virus == 1) # filters out PBS 
data_FHV_PBS <- dplyr::filter(data_FHV, Virus == 0) # Filters out virus

# Checking data
head(data_FHV)
head(data_FHV_noPBS)
head(data_FHV_PBS)

# Checking factors
levels(data_FHV$Wol)
levels(data_FHV$Virus)

```

# Coxme Analysis 
Applying cox-me model to determine the impact of Wolbachia on survival when flies are infected with FHV.

* Wolbachia = predictor
* biol and tech replicates = random effects

```{r Coxme fitting}
# Fitting the model
fit1 <- coxme(Surv(Days, Survival) ~ Wol + (1 | BiolRep/TechRep), data = data_FHV_noPBS)
```

# Assumptions 
Testing the assumptions of coxme
1. Independent of treatments, virus impacts survival 
2. Independent of viral infection, Wolbachia doesn't impact survival (PBS)
3. Proportional hazards 

Results:
1. Passed: p-value for virus coef is significant 
2. Passed: p-value for wol coef is NOT significant
3. Passed: Global and wol coef passed p-value is NOT significant

Conclusions: 
Assumptions are met cox-me is appropriate analysis 



```{r Coxme Assumptions}
# 1. Independent of treatments virus impacts survival 
fitSus <- coxme(Surv(Days, Survival) ~ Virus + (1 | BiolRep/TechRep), data= data_FHV) 
fitSus 

# 2. Independent of viral infection Wolbachia doesn't impact survival (PBS)
fitPBS <- coxme(Surv(Days, Survival) ~ Wol + (1 | BiolRep/TechRep), data = data_FHV_PBS) 
fitPBS

# 3. Proportional hazards: Visualise
# Plot
LMLplot <- plot(survfit(Surv(Days, Survival) ~ Wol, data = data_FHV_noPBS), fun = "cloglog")

# 3. Proportional hazards: Formally test 
ftest <- cox.zph(fit1)
print(ftest) # prints test results
ggcoxzph(ftest) # plots test results
```


# Conclusions 
Wolbachia impacts survival coxme
* The HR of flies with wol is 0.005933289 
* For every fly without wolbachia 0.005933289 flies with Wolbachia die. 
* For each time point, flies with wolbachia are ~99% (1-0.08902321 = 0.9940667) less likely to die from virus than flies without Wolbachia

# Summary of analysis
```{r Summary Coxme output}
# Extract the HR estimates and confidence intervals
coef_cox<- as.data.frame(summary(fit1)$coefficients) # extract coef
HR_cox <- as.data.frame(exp(coef_cox)) # convert coefficient to HR 
CI_cox <- as.data.frame(exp(confint(fit1))) # CIs for HR 

# Extracts p-value 
extract_coxme_table <- function (mod){
  beta <- mod$coefficients 
  nvar <- length(beta)
  nfrail <- nrow(mod$var) - nvar
  se <- sqrt(diag(mod$var)[nfrail + 1:nvar])
  z<- round(beta/se, 2)
  p<- signif(1 - pchisq((beta/se)^2, 1), 2)
  table=data.frame(cbind(beta,se,z,p))
  return(table)
}
extract_coxme_table(fit1)
p_cox <- as.data.frame(extract_coxme_table(fit1)$p)

# Create data-frame
df_cox <- bind_cols(coef_cox, HR_cox, CI_cox, p_cox) %>% 
  rename(coef = `summary(fit1)$coefficients...1`, 
         HR = `summary(fit1)$coefficients...2`, 
         "LB_2.5%" = "2.5 %",
         "UB_97.5%" = "97.5 %", 
         pval = "extract_coxme_table(fit1)$p") %>% 
  mutate(test = "coxme") %>% 
  tibble::rownames_to_column("factor")

# confirms extraction 
print(fit1)
View(df_cox)
```

```{r Visualising Coxme output}
# Graph of HR and CI 
plot_coxme <- ggplot(df_cox, aes(x = factor, y = HR, colour = test, group_by(test))) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = `LB_2.5%`, ymax = `UB_97.5%`, color = test), width = 0.1) +
  coord_flip() +
  xlab("") +
  ylab("Hazard Ratio") +
  scale_color_manual(values = c("blue")) +
  theme_minimal()
```

# Graphing Kaplan-Meier survival curves 

The curves generated are Kaplan-Meier but the analysis that has been conducted is coxme

```{r Creating Kaplan-Meier curves}
# Creating the survival curve 
curves <- survfit(Surv(Days, Survival) ~ Wol, data=data_FHV_noPBS)

curves_PBS <- survfit(Surv(Days, Survival) ~ Wol+ Virus, data=data_FHV)
```

```{r Plotting Kaplan-Meier curves}
# Without PBS
graph_FHV <- ggsurvplot(curves,
           data = data_FHV_noPBS,              
           axes.offset = FALSE,     
           line = c(1,1,1,2,2,2),
           xlim = c(0,35),
           ylim = c(0,1.01),
           xlab = "Days Post Infection",
           ylab = "Proportional survival",
           legend.title=element_blank(), 
           break.time.by = 2, 
           break.y.by = 0.2,
           conf.int = TRUE,
           ggtheme = theme_classic2(),
           palette = c("black", "red"),
           legend = "none"
) |> ggpar(
  font.x        = c(30, "bold", "black"),          
  font.y        = c(30, "bold", "black"),      
  font.xtickslab = c(30, "plain", "black"),
  font.ytickslab = c(30, "plain", "black")
) 

# With PBS
graph_FHV_PBS <- ggsurvplot(curves_PBS,
           data = data_FHV,         
           axes.offset = FALSE,            #changes axes to start at zero
           line = c(2,1,2,1),
           xlim = c(0,35),
           ylim = c(0,1.01),
           xlab = "Days Post Infection",
           ylab = "Proportional survival",
           legend.title=element_blank(), 
           break.time.by = 2, 
           break.y.by = 0.2,
           conf.int = TRUE,
           ggtheme = theme_classic2(),
           palette = c("black", "black", "red", "red"),
           legend = "none"
) |> ggpar(
  font.x        = c(30, "bold", "black"),          
  font.y        = c(30, "bold", "black"),      
  font.xtickslab = c(30, "plain", "black"),
  font.ytickslab = c(30, "plain", "black")
)
```

# Exporting outputs 
Run outside of rmd
```{r Exporting outputs}

# Saving coxme summary
write.csv(df_cox, file = "Survival_FHV_Figure_1_C_output_coxme_HR.csv", row.names = FALSE)

# Exporting stats output
sink("Survival_FHV_Figure_1_C_output_coxme_summary.txt") # change file name
summary(fit1) # prints coxme output to txt file UP TO HERE
sink() # closes the connection **CRITICAL**


# Saving coxme plot 
ggsave("Survival_FHV_Figure_1_C_output_plot_coxme.pdf", 
       plot = plot_coxme, 
       height = 7.5, 
       width = 15) 

# Saving Kaplan-Meier survival curves 
ggsave("Survival_FHV_Figure_1_C_output_plot_survival_curves.pdf", 
       plot = graph_FHV$plot, 
       height = 7.5, 
       width = 15)

ggsave("Survival_FHV_Figure_2_C_output_plot_survival_curves_PBS.pdf", 
       plot = graph_FHV_PBS$plot, 
       height = 7.5, 
       width = 15)
```