---
title: "Figure 3: Infectivity of DCV and FHV"
date: "2023-09-18"
output: html_document
---


## Experiment Description 
Virus: Progney DCV and FHV (W- and W+)
This document contains the analysis comparing the infectivity of virus from Drosophila with and without Wolbachia
Overall to measure the impact of Wolbachia on the infectivity of DCV and FHV. Infectivity was measured as the number of infectious virus /total virus 


## Approach
Steps
 * Transformation (log scale) DONE
 * t-test DONE
 * GLM and Assumptions DONE
 * Compare outputs DONE 
 * Presentation DONE
 
# Preparation 
 
```{r Preparation loading libraries}
library(tidyverse)
library(ggpubr)
library(moments)
library(car)
library(rstatix)
```

## Importing Data 
sample        = Sample number
IU_mL         = Measure of concentration of infectious particles (IU/ml)
treatment     = Progeny virus from drosophila with (W+) or without (W-) Wolbachia
virus         = The virus (DCV or FHV)
biol_rep      = The biological replicate
total_virus   = Concentration of total virus (Total/ml)

```{r Importing data}
# Importing measures of infectious particles and total particles
data <- as.data.frame(read_csv("Infectivity_DCV_FHV_Figure_3_data.csv"))
```


## Transformation 
Transform (Log10) the data because it is not normally distributed. 
Transform the total and infectious virus before taking the ratio. 
Infectivity = log10(IU)/log10(Total)
NB 
While the measures are concentrations if we assume we are looking at the number of virus particles that are infectious in 1 ml then it simplifies
logIU/ml x 1 ml = logIU
logtotal/mL x 1ml = logtotal. 
infectivity = logIU/logtotal 

```{r Transformation}
# log10 transformation
data <- data %>% 
  mutate(log_total_virus = log10(total_virus)) %>% # log10 transform total virus
  mutate(log_IU_ml = log10(IU_mL)) %>% # log10 transform IU 
  mutate(log_infectivity_ratio = log_IU_ml/log_total_virus) %>% # calculate log infectivity
  mutate(infectivity_ratio = IU_mL/total_virus) %>%  # calculate untransformed infectivity
  mutate(specific_infectivity = total_virus/IU_mL)  # Calculates specific infectivity (number of total to infectious particles)

# Collapsing the dataframe so there is only one point for each biol rep that is the average of two tech reps 

# Averaging the technical replicates
data_1 <- data %>%
  group_by(biol_rep) %>%
  mutate(average_log_infectivity_ratio = mean(log_infectivity_ratio)) %>% 
  mutate(average_infectivity_ratio = mean(infectivity_ratio)) %>% 
  mutate(average_specific_infectivity = mean(specific_infectivity)) %>% 
  select(treatment, 
         virus, 
         biol_rep,
         average_infectivity_ratio, 
         average_log_infectivity_ratio, 
         average_specific_infectivity) # selecting only the average data 

# Collapsing so there is only the average for each biol rep 
c <- seq(from = 1, by =2, length.out = 12)
data_2 <- data_1[c,]

# Formatting
data_2 <- data_2 %>%
  mutate(average_log_infectivity_ratio = as.double(average_log_infectivity_ratio)) %>%
  mutate(average_infectivity_ratio = as.double(average_infectivity_ratio)) %>%
  mutate(average_specific_infectivity = as.double(average_specific_infectivity)) %>%
  mutate(treatment = as.factor(treatment)) %>%
  mutate(virus = as.factor(virus))

# Changing the names of the treatments 
# Changing it so that it is W- for Tet and W+ for Wol 
data_2$treatment <- gsub("tet", "W-", data_2$treatment)
data_2$treatment <- gsub("wol", "W+", data_2$treatment)
data_2$treatment <- factor(data_2$treatment, levels = c("W-", "W+"))
data_2$virus <- factor(data_2$virus, levels = c("DCV", "FHV"))

# Subset data based on the virus 
data_DCV <- data_2 %>% 
  subset(virus == "DCV")

data_FHV <- data_2 %>% 
  subset(virus == "FHV")

# Subset the data based on the Wolbachia status 

data_DCV_no_Wol <- data_DCV %>% 
  subset(treatment == "W-")
  
data_DCV_Wol <- data_DCV %>% 
  subset(treatment == "W+")

data_FHV_no_Wol <- data_FHV %>% 
  subset(treatment == "W-")
  
data_FHV_Wol <- data_FHV %>% 
  subset(treatment == "W+")
```

The specific infectivity is essentially the inverse of the measure of infectivity we did 
specific infectivity = total particles/infectious 

## Specific Infectivity

```{r Calculating the specific infectivity as measured in the literature}
# Calculating the average spec_infectivity for each group 
DCV_no_Wol_spec_infect <- mean(data_DCV_no_Wol$average_specific_infectivity)
DCV_Wol_spec_infect <- mean(data_DCV_Wol$average_specific_infectivity)
FHV_no_Wol_spec_infect <- mean(data_FHV_no_Wol$average_specific_infectivity)
FHV_Wol_spec_infect <- mean(data_FHV_Wol$average_specific_infectivity)

av <- mean(data_DCV_no_Wol$average_specific_infectivity)

# Creating a table with these results 
total_to_IU <-
  rbind(
    DCV_no_Wol_spec_infect,
    DCV_Wol_spec_infect,
    FHV_no_Wol_spec_infect,
    FHV_Wol_spec_infect
  )
total_to_IU <- round(total_to_IU, digits = -2)

Virus <-  c("DCV W-", "DCV W+", "FHV W-", "FHV W+")

spec_infect_number <- as.data.frame(total_to_IU) 
spec_infect_number <- spec_infect_number %>% 
  rename( "Total Virus to 1 spec_infectious particle" = "V1") %>% 
  tibble::rownames_to_column( "Delete") %>% 
  cbind(Virus) %>% 
  select(Virus, "Total Virus to 1 spec_infectious particle", -Delete)
```

## T-test
There is no statistical difference when comparing W- and W+ virus for the same virus. 

```{r T-test}
#DCV 
t_test_DCV <- t.test(data = data_DCV, 
                 average_log_infectivity_ratio ~ treatment)
t_test_DCV

#FHV 
t_test_FHV<- t.test(data = data_FHV, 
                 average_log_infectivity_ratio ~ treatment)
t_test_FHV
```

## GLM 

Results 
* model 2 best fit by AIC although not much deviance (no interaction)
* 
 
 
```{r GLM }
## GLM ## 
# With interaction 
model_1 <- glm(formula = average_log_infectivity_ratio ~ treatment * virus,
             family = binomial, 
             data = data_2)

# Without interaction
model_2 <- glm(formula = average_log_infectivity_ratio ~ treatment + virus,
             family = binomial, 
             data = data_2)

# Best model AIC 
AIC(model_1, model_2)
AIC(model_2, model_3)

# Comparing models ANOVA
anova(model_1, model_2)

# Summary of the model
summary(model_2)
```

# GLM Assumptions

1. Homogeneity of variances (PASS)
2. Normality (PASS)
3. Outliers (PASS)


```{r GLM Assessing Assumptions}

# 1. Homogeneity of variances
plot(model_2, 1)
leveneTest(average_log_infectivity_ratio ~ 
             virus, 
           data = data_2)

# 2. Normality
plot(model_2, 2)
shapiro_test(residuals(model_2))

# 3. Outliers 
plot(model_2, 1)
outlierTest(model_2)
```


# Presenting the results 

* error bars = SEM 
* bar is the mean 

```{r Presenting the results}

summary_data <- data_2 %>%
  group_by(treatment, virus) %>%
  summarize(
    Mean = mean(average_infectivity_ratio),
    SE = sd(average_infectivity_ratio) / sqrt(n())
  )

# Graphing the data 
infectivity <-  ggplot(data_2) +
  geom_jitter(
    aes(
      x = interaction(treatment, virus),
      y = average_infectivity_ratio, # not transformed
      col = treatment
    ),
    size = 5,
    position = position_jitter(width = 0.1, height = 0.0001),
    show.legend = FALSE
  )  +
  scale_color_manual(values = c("black", "red")) +
  scale_shape_manual(values = c(16, 1)) +
  stat_summary(
    data = summary_data,
    aes(
      x = interaction(treatment, virus),
      y = Mean,
      group = virus
    ),
    fun.data = mean_se,
    geom = "crossbar",
    width = 0.2,
    # Adjust the width of the bars
    col = c("black", "red", "black", "red")  # Color of the bars
  ) +
  # Adjust the appearance of summary bars and error bars
  geom_errorbar(
    data = summary_data,
    aes(
      x = interaction(treatment, virus),
      ymin = Mean - SE,
      ymax = Mean + SE,
      group = virus
    ),
    width = 0.1,
    # Width of error bars
    col = c("black", "black", "red", "red")  # Color of error bars
  ) +
  theme_classic() +
  labs(x = "",
       y = expression(bold("Infectivity ratio"))) +  # Use "\n" for line break
  theme(
    plot.margin = unit(c(3, 1, 1, 2), "lines"),
    # adjust plot margins
    axis.title = element_text(
      size = 30,
      face = "bold",
      colour = "black"
    ),
    # center justify axis title text
    axis.title.y = element_text(
      size = 30,
      face = "bold",
      colour = "black",
      hjust = 0.5,
      vjust = 0.1
    ),
    # center text on the Y-axis
    axis.text = element_text(
      size = 30,
      face = "bold",
      colour = "black"
    ),
    # increase size of axis tick labels
    plot.title = element_text(size = 30)
  ) + # increase size of plot title text
  coord_cartesian(ylim = c(0, 0.005)) + # sets the maximum y-axis
  scale_x_discrete(labels = c("W-\n DCV", "W+\n DCV", "W-\n FHV", "W+\n FHV")) +
  scale_y_continuous(breaks = c(0, 0.0025, 0.005))

```


```{r Export plots run outside of RMD}
# Exporting the specific infectivity summary table
write.csv(spec_infect_number, file = "Infectivity_DCV_FHV_Figure_3_Output_Specific_Infectivity.csv")

# Exports T-test output 
# DCV 
sink("Infectivity_DCV_FHV_Figure_3_Output_DCV_T_test.txt") 
print(t_test_DCV) 
sink() # closes the connection **CRITICAL**

# FHV
sink("Infectivity_DCV_FHV_Figure_3_Output_FHV_T_test.txt") 
print(t_test_FHV) 
sink() # closes the connection **CRITICAL**

# Exports GLM
# DCV and FHV model 
sink("Infectivity_DCV_FHV_Figure_3_Output_GLM.txt") # change file name
print(model_2) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**

# DCV and FHV model SUMMARY 
sink("Infectivity_DCV_FHV_Figure_3_Output_GLM_summary.txt") # change file name
print(summary(model_2)) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**

# Exporting figure
ggsave(filename = "Infectivity_DCV_FHV_Figure_3_output_plot.pdf",
       plot = infectivity, 
       width = 8, 
       height = 4)
```

