---
title: "Accumulation of FHV in flies with and without Wolbachia"
author: "Angelique Asselin"
date: "2023-09-19"
output: html_document
---
# Experimental details
 Virus:     FHV
 Titre:     10E09 IU/mL 
 Fly:       w1118 wMelPop and w1118 wMelPop-Tet
 Figure:    1D
 
# Rationale 
To determine if Wolbachia has an impact on measured FHV levels three hours post infection. 

# Approach
Welch T-test 

# Preparation 
```{r Preparation: Loading library}
library(car)
library(rstatix)
library(tidyverse)
library(ggpubr)
library(dixonTest)
```

# Importing Ct data 

Data
Target_takeoff    = The ct value for reactions targeting FHV 
Reference_takeoff = The ct value for reactions targeting RpL32
Sample_number     = The number of the sample note that samples have two technical replicates 

Data_ID 
Contains information such as the time of collection (Time) for each sample the replicate (Replicate) number(n=6) and the fly line the data are from (Line; 0 = w1118-wMelpopTet, 1= w1118-wMelPop)

```{r Importing Ct Data}
# Importing raw Ct values
data_FHV_acc<- as.data.frame(read_csv("Accumulation_FHV_Figure_1_D_data.csv"))

# Importing sample ID information
data_ID <- as.data.frame(read_csv("Accumulation_FHV_Figure_1_D_data_sample_ID.csv"))
```

# Calculating MNE from Ct Data 
This analysis utilises Equation (2) and (3) from 
DOI: 10.1093/bioinformatics/btg157

```{r Analysing Ct Data}
# Setting the data
Data <- data_FHV_acc

#Sets % SEM cutoff for samples (default 20%)
SEMCutoff = 20

# Do you wish to automatically censor data over the above SEM% threshold? 0 = no, 1 = yes (default yes)
UseCutoff = 1

# Change std Amp. efficiency if necessary
TarAmp = 1.99 # FHV primer efficiency derived previously in the lab 
RefAmp = 1.64 # rpl32 primer efficiency derived previously in the lab 

# Sorting the data for analysis

Normalised <-
  data.frame("Sample", "MNE", "SE of MNE", "SE of MNE as % (x<=20%)")
i = 0
val = 1:nrow(Data)
val <- val[seq(1, (length(val)), 2)]
w = unlist(c(Data[4]))
x = unlist(c(Data[3]))
y = unlist(c(Data[2]))
s = matrix(x, nrow = 1, ncol = nrow(Data))
v = matrix(w, nrow = 1, ncol = nrow(Data))
t = matrix(y, nrow = nrow(Data), ncol = 1)
list <-
  data.frame("Censored Samples", "MNE", "SE of MNE", "SE of MNE as % (x>25%)")

# For loop processes all the data and adds it to the dataframe
for (i in val) {
  # Calculates MNE of Duplicates
  MNE2 = (RefAmp ^ ((v[1, i] + v[1, i + 1]) / 2)) / (TarAmp ^ ((s[1, i] +
                                                                  s[1, i + 1]) / 2))
  
  # Calculates SE of MNE
  SEofCTTAR = sd(c(s[1, i], s[1, i + 1])) / sqrt(2)
  SEofCTREF = sd(c(v[1, i], v[1, i + 1])) / sqrt(2)
  
  SEofMNE = MNE2 * (((log(TarAmp) * SEofCTTAR) ^ 2) + ((log(RefAmp) * SEofCTREF) ^
                                                         2)) ^ .5
  
  # Calculates SE of MNE as %
  SEofMNEasP = (SEofMNE / MNE2) * 100
  
  # Adds each set of data to the data frame (you may wish to change the digits= option to be smaller if you wish)
  if (UseCutoff == 1) {
    if (SEofMNEasP > SEMCutoff) {
      cat(
        "Sample:",
        t[i, 1],
        "was censored due to exceeding SEM cutoff of",
        SEMCutoff,
        "with a value of",
        (round(SEofMNEasP, digits = 2)),
        "\n"
      )
      list[nrow(list) + 1,] = c(
        t[i, 1],
        round(MNE2, digits = 10),
        round(SEofMNE, digits = 5),
        round(SEofMNEasP, digits = 2)
      )
    }
    else {
      Normalised[nrow(Normalised) + 1,] = c(
        t[i, 1],
        round(MNE2, digits = 10),
        round(SEofMNE, digits = 5),
        round(SEofMNEasP, digits = 2)
      )
    }
  }
  else {
    Normalised[nrow(Normalised) + 1,] = c(
      t[i, 1],
      round(MNE2, digits = 10),
      round(SEofMNE, digits = 5),
      round(SEofMNEasP, digits = 2)
    )
  }
}
```

```{r Formats MNE output}
# Formats the analysed dataset
MNE <- Normalised %>%
  rename(
    Sample = X.Sample.,
    FHV_MNE = X.MNE.,
    FHV_SE_MNE = X.SE.of.MNE.,
    FHV_SE_Percent = X.SE.of.MNE.as....x..20...
  ) %>%
  slice(-1) %>% 
  select(Sample, everything())
```

# Formating data
Joins the data with the associated Data_ID information
* Line (flies with or without Wolbachia) 
* Time 
* Replicate 

```{r Sample information}
# Code the samples as the same 
MNE$Sample <- as.double(MNE$Sample)
data_ID$Sample <- as.double(data_ID$Sample)

# Joining the MNE dataframe with the sample ID dataframe 
data_MNE <- full_join(data_ID, MNE, by = join_by(Sample))
```

# Formating MNE for analysis and graphing 

```{r Formating MNE for analysis and graphing}
# Coding the factors correctly 
data_MNE$Time <- as.factor(data_MNE$Time)
data_MNE$FHV_MNE <- as.numeric(data_MNE$FHV_MNE)
data_MNE$Line <- as.factor(data_MNE$Line)

# Log-transforming data
data_MNE <- data_MNE %>% 
  mutate(log_10_FHV_MNE = log10(FHV_MNE)) 

```

# Graphing 
```{r Graphing}
FHV <-  ggplot(data_MNE, aes(x = interaction(Line,Time), y = log_10_FHV_MNE))+ 
  geom_point(aes(col = Line), size = 6, position = position_jitter(width = 0.01)) +
  theme_classic()+ 
  stat_summary(fun.data = mean_se, 
               geom = "errorbar", 
               size = 1,
               width = 0.3, 
               aes(group = interaction(Line, Time), col = Line)) +
  stat_summary(fun = mean,
               geom = "crossbar",
               width = 0.15,
               position = position_dodge(width = 0.3), 
               aes(group = interaction(Line, Time), col = Line)) +
  labs(x = NULL, y = expression(bold("Relative abundance of FHV RNA ")), color = "Group") +
  scale_color_manual(values = c("black", "red")) +
  theme(plot.margin = unit(c(3, 1, 1, 2), "lines"),
        axis.text.x = element_text(face= "bold", size = 30, colour = "black", hjust = 0.5, vjust = 0.1),  # Remove x-axis text
        axis.text.y = element_text(face= "bold", size = 30, colour = "black", hjust = 0.5, vjust = 0.1), 
        axis.title = element_text(size = 30))+
  scale_x_discrete(labels = c("W-\n                          0 hours", "W+\n", "W-\n                              72 hours ", "W+\n")) +
  guides(color = FALSE)

```

# Analysis 

Comparing the virus levels of two groups. 
Flies with Wolbachia and flies without Wolbachia. 
Compare each time point separately. 

```{r FHV T-test}
# Comparing 0 hpi ##
data_MNE_sub_0 <- subset(data_MNE, Time == "0")

# Removing outliers (#sample 11)
data_MNE_sub_0_outlier <- subset(data_MNE_sub_0, Sample != 11)

# T-test
FHV_t_test_0 <- t.test(log_10_FHV_MNE ~ Line, data = data_MNE_sub_0)
FHV_t_test_0
print(FHV_t_test_0)

# T-test without outlier 
FHV_t_test_out_0 <- t.test(log_10_FHV_MNE ~ Line, data = data_MNE_sub_0_outlier)
FHV_t_test_out_0

## Comparing 3 hpi ##
 
data_MNE_sub_3 <- subset(data_MNE, Time == "3")

# Removing outliers (#sample 11)
data_MNE_sub_3_outlier <- subset(data_MNE_sub_3, Sample != 11)

# T-test
FHV_t_test_3 <- t.test(log_10_FHV_MNE ~ Line, data = data_MNE_sub_3)
FHV_t_test_3
print(FHV_t_test_3)

# T-test without outlier 
FHV_t_test_out_3 <- t.test(log_10_FHV_MNE ~ Line, data = data_MNE_sub_3_outlier)
FHV_t_test_out_3

``` 

# Results from analysis 

0 hours post infection 
* Wolbachia has no impact on the relative amount of FHV at 0 hpi (therefore the innoculumn was equivalent)


3 hours post infection 
* Wolbachia has an impact on the relative abundance of FHV after 3 hour post infection. 
* outlier doesn't impact overall conclusion therefore keep the outlier in for the analysis as there is no evidene compelling it's removal 


# Assumptions 
1. Independence = PASS (in the design)
2. Normality of dependent variable = PASS (Fit in the confidence interval)
3. No significant outliers = FAIL (outlier in the Wolbachia 3 hpi data # sample 11, however no impact of removing this outlier on the overall conclusions of this analysis) 
4. Dependent variable is continuous (PASS)

OUTLIER
There looks like there is an outlier in the Wolbachia samples but when I checked the time 0 data it was not detectable in this replicate at that time point suggesting that the virus has accumulated so there was a productive infection it just didn't replicate to as high a levels. Because there was an outlier in the data wanted to see if it impacted the outcome. There is still a measured statistical difference regardless of whether the outlier was included but the variation and error is greater for the data when including the outlier. 

```{r Assumptions}
## 0 hpi ## 
# 2. Normality 
FHV_0_0hpi <- data_MNE_sub_0 %>% 
  subset(Line == "0")

FHV_1_0hpi <- data_MNE_sub_0 %>% 
  subset(Line == "1")

ggqqplot(FHV_0_0hpi$log_10_FHV_MNE) 
ggqqplot(FHV_1_0hpi$log_10_FHV_MNE) 

# 3. Outliers 
# In the data from Flies without Wolbachia
dixonTest(FHV_0_0hpi$log_10_FHV_MNE)

# In data from flies with Wolbachia 
dixonTest(FHV_1_0hpi$log_10_FHV_MNE) #None 


## 3 hpi ## 
# 2. Normality 
FHV_0_3hpi <- data_MNE_sub_3 %>% 
  subset(Line == "0")

FHV_1_3hpi <- data_MNE_sub_3 %>% 
  subset(Line == "1")

ggqqplot(FHV_0_3hpi$log_10_FHV_MNE) 
ggqqplot(FHV_1_3hpi$log_10_FHV_MNE) 

# 3. Outliers 
# In the data from Flies without Wolbachia
dixonTest(FHV_0_3hpi$log_10_FHV_MNE)

# In data from flies with Wolbachia 
dixonTest(FHV_1_3hpi$log_10_FHV_MNE) # Sample 11 Value -1.15 

```

# Exporting 
```{r Exporting Graph}
name <- paste("Figure", "1", "D", sep = "_")
pdf_name <- paste(name, "pdf", sep = ".")
ggsave(filename = pdf_name,
       plot = FHV, 
       width = 15, 
       height = 7.5)
```

```{r FHV T-test Output RUN OUT OF RMD}
# 0 hpi 
sink("Accumulation_FHV_Figure_1_D_stats_output_0hpi.txt") # add date and data_IFB_01 name | creates the txt file in WD
print(FHV_t_test_0) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**

# 3 hpi 
sink("Accumulation_FHV_Figure_1_D_stats_output_3hpi.txt") # add date and data_IFB_01 name | creates the txt file in WD
print(FHV_t_test_3) # prints coxme output to txt file
sink() # closes the connection **CRITICAL**
```
